### 2021-10-11 papers/thesis/resubmitted/ch3-disk-slices/disk-slices.tex
Made a portion more consistent with rest of thesis
Fixed typos for Jac mats defs. Added brackets to eqn numbers in citations

Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$. Whereas the sparsity pattern and entries derived for equations on the triangle \cite{olver2018recurrence, olver2019triangle} and for equations on the disk \cite{vasil2016tensor} with the frameworks discussed earlier both result from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries. In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This compares favourably to $O(p^6)$ operations if one proceeds na\"ively. Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.

@@

\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then, for $n = 0,1,2,\dots$, $k = 0,\dots,n$, the 2D polynomials defined on $\Omega$ given by
\bseqn{
	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)})
}
are orthogonal polynomials with respect to the weight $W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)})$ on $\Omega$. 
\end{proposition}

@@

Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
\bseqn{
	J_{x/y}^{(a,b,c,d)} &= 
		\begin{pmatrix}
			B^{x/y}_0 & A^{x/y}_0 & & & & \\
			C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
			& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
			& & C^{x/y}_3 & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
For $J_x^{(a,b,c,d)}$, the blocks are diagonal (sub-block-bandwidths $(0,0)$) for $n = 0,1,2,\dots$:
\bseqn{
	A^x_n &:= 
		\begin{pmatrix}
			\alphaabcd_{n+1,0,1} & 0 & \hdots & 0 \\
			& \ddots & & \vdots & \\
			& & \alphaabcd_{n+1,n,1} & 0 \\
		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
	B^x_n &:= 
		\begin{pmatrix}
			\alphaabcd_{n,0,2} & & \\
			& \ddots & \\
			& & \alphaabcd_{n,n,2} \\
		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
	C^x_n &:= \big( A^x_{n-1} \big)^\top \in \R^{(n+1)\times n}, \quad (n \ne 0)
}
and for $J_y^{(a,b,c,d)}$ the blocks are tridiagonal (sub-block-bandwidths $(1,1)$):
\bseqn{
	A^y_n &:= 
		\begin{pmatrix}
			\betaabcd_{n,0,8} & \betaabcd_{n,0,9} & & & \\
			\betaabcd_{n,1,7}& \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & \betaabcd_{n,n,7} & \betaabcd_{n,n,8} & \betaabcd_{n,n,9} \\
		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
	B^y_n &:= 
		\begin{pmatrix}
			\betaabcd_{n,0,5} & \betaabcd_{n,0,6} & & \\
			\betaabcd_{n,1,4} & \ddots & \ddots & \\
			& \ddots & \ddots & \betaabcd_{n,n-1,6} \\
			& & \betaabcd_{n,n,4} & \betaabcd_{n,n,5}
		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
	C^y_n &:= 
		\begin{pmatrix}
			\betaabcd_{n,0,2} & \betaabcd_{n,0,3} & & \\
			\betaabcd_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots & \betaabcd_{n,n-2,3} \\
			& & \ddots & \betaabcd_{n,n-1,2} \\
			& & & \betaabcd_{n,n,1} \\
		\end{pmatrix} \in \R^{(n+1)\times n}, \quad (n \ne 0).
}
Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs, meaning that the sparsity is not limited to the specific disk-slice case.

@@

We can combine each system in \bsrefeqns{eqn:ds:jacobimatricesdefinitionx}{eqn:ds:jacobimatricesdefinitiony} into a block-tridiagonal system:
\bseqn{
	\renewcommand\arraystretch{1.3}
	&\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\bighdopabcd(x,y) \\
	&=
	\begin{pmatrix}
		\hdopabcd_{0,0} & 0 & 0 & 0 & \hdots  \\
	\end{pmatrix}^\top,
}
where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
\bseqnnumber{
A_n &:= \begin{pmatrix}
		A^x_n \\
		A^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)}, \quad
C_n := \begin{pmatrix}
		C^x_n \\
		C^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \label{eqn:ds:jacobisubmats1} \\
B_n &:= \begin{pmatrix}
		B^x_n \\
		B^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y) := \begin{pmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}. \label{eqn:ds:jacobisubmats2}
}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+2}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bighdopabcd_{-1}(x,y) := 0 \\
		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y)
	\end{cases}.
}
Note that we can define an explicit $\Dnt$ as follows:
\bseqnnumber{
\Dnt := \begin{pmatrix}
		\frac{1}{\alphaabcd_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaabcd_{n+1,n,1}} & \\
		\eta_{0} & 0 & \hdots & 0 & \eta_{1} & \hdots & \hdots & \eta_{n+1}
	    \end{pmatrix}, \label{eqn:ds:Dtdef}
}
 where
 \bseqn{
	\eta_{n+1} &= \frac{1}{\betaabcd_{n,n,9}}, \\ 
	\eta_{n} &= -\frac{1}{\betaabcd_{n,n-1,9}} \big( \betaabcd_{n,n,8} \: \eta_{n+1} \big), \\
	\eta_j &= - \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
	\eta_0 &= - \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
}
It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopabcd_{0}(x,y)$  through $\bighdopabcd_{n}(x,y)$ in optimal $O(n^2)$ complexity.

@@

It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in \bsrefeqn{eqn:ds:Hrecurrence}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}. We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case. We thus first need to define a new set of `interim' 1D OPs.
\begin{definition}\label{def:ds:InterimOPconstruction}
Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
\bseqnnumber{
	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ds:ipgenjact}
}
where
\bseqnnumber{
	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
}
Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjact}.
\end{definition}
Note that the OPs $\{\genjac_n^{(a,b,2c)}\}$ are then equivalent to the OPs $\{\genjact_n^{(a,b,c,c)}\}$. Let the recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d)}\}$ be given by:
\bseqnnumber{
	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
}
\begin{proposition}
There exist constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ such that
\bseqnnumber{
	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansioni} \\
	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansionii}
}
\end{proposition}

>>>>

Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$. Whereas the sparsity pattern and entries derived for equations on the triangle \cite{olver2018recurrence, olver2019triangle} and for equations on the disk \cite{vasil2016tensor} with the frameworks discussed earlier both result from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[(18.2.12)]{DLMF} and quadrature rules to determine the entries. In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This compares favourably to $O(p^6)$ operations if one proceeds na\"ively. Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.

@@

\begin{proposition}[{\cite[p.55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then, for $n = 0,1,2,\dots$, $k = 0,\dots,n$, the 2D polynomials defined on $\Omega$ given by
\bseqn{
	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)})
}
are orthogonal polynomials with respect to the weight $W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)})$ on $\Omega$. 
\end{proposition}

@@

Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
\bseqn{
	J_{x/y}^{(a,b,c,d)} &= 
		\begin{pmatrix}
			B_{x/y,0} & A_{x/y,0} & & & & \\
			C_{x/y,1} & B_{x/y,1} & A_{x/y,1} & & & \\
			& C_{x/y,2} & B_{x/y,2} & A_{x/y,2}  & & & \\
			& & C_{x/y,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
For $J_x^{(a,b,c,d)}$, the blocks are diagonal (sub-block-bandwidths $(0,0)$) for $n = 0,1,2,\dots$:
\bseqn{
	A_{x,n} &:= 
		\begin{pmatrix}
			\alphaabcd_{n+1,0,1} & 0 & \hdots & 0 \\
			& \ddots & & \vdots & \\
			& & \alphaabcd_{n+1,n,1} & 0 \\
		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
	B_{x,n} &:= 
		\begin{pmatrix}
			\alphaabcd_{n,0,2} & & \\
			& \ddots & \\
			& & \alphaabcd_{n,n,2} \\
		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
	C_{x,n} &:= \big( A^x_{n-1} \big)^\top \in \R^{(n+1)\times n}, \quad (n \ne 0)
}
and for $J_y^{(a,b,c,d)}$ the blocks are tridiagonal (sub-block-bandwidths $(1,1)$):
\bseqn{
	A_{y,n} &:= 
		\begin{pmatrix}
			\betaabcd_{n,0,8} & \betaabcd_{n,0,9} & & & \\
			\betaabcd_{n,1,7}& \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & \betaabcd_{n,n,7} & \betaabcd_{n,n,8} & \betaabcd_{n,n,9} \\
		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
	B_{y,n} &:= 
		\begin{pmatrix}
			\betaabcd_{n,0,5} & \betaabcd_{n,0,6} & & \\
			\betaabcd_{n,1,4} & \ddots & \ddots & \\
			& \ddots & \ddots & \betaabcd_{n,n-1,6} \\
			& & \betaabcd_{n,n,4} & \betaabcd_{n,n,5}
		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
	C_{y,n} &:= 
		\begin{pmatrix}
			\betaabcd_{n,0,2} & \betaabcd_{n,0,3} & & \\
			\betaabcd_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots & \betaabcd_{n,n-2,3} \\
			& & \ddots & \betaabcd_{n,n-1,2} \\
			& & & \betaabcd_{n,n,1} \\
		\end{pmatrix} \in \R^{(n+1)\times n}, \quad (n \ne 0).
}
Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs, meaning that the sparsity is not limited to the specific disk-slice case.

@@

We can combine each system in \bsrefeqns{eqn:ds:jacobimatricesdefinitionx}{eqn:ds:jacobimatricesdefinitiony} into a block-tridiagonal system:
\bseqn{
	\renewcommand\arraystretch{1.3}
	&\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\bighdopabcd(x,y) \\
	&=
	\begin{pmatrix}
		\hdopabcd_{0,0} & 0 & 0 & 0 & \hdots  \\
	\end{pmatrix}^\top,
}
where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
\bseqnnumber{
	A_n &:= 
		\begin{pmatrix}
                    	A_{x,n} \\
			A_{y,n}
		\end{pmatrix}, \quad \in \R^{2(n+1)\times(n+2)}, \quad
	C_n &:= 
		\begin{pmatrix}
                    	C_{x,n} \\
			C_{y,n}
		\end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \label{eqn:ds:jacobisubmats1} \\
	B_n &:= 
		\begin{pmatrix}
                    	B_{x,n} \\
			B_{y,n}
		\end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
	G_n(x,y) := 
		\begin{pmatrix}
			xI_{n+1} \\
			yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}. \label{eqn:ds:jacobisubmats2}
}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+2}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p.78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bighdopabcd_{-1}(x,y) := 0 \\
		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y)
	\end{cases}.
}
Note that we can define an explicit $\Dnt$ as follows:
\bseqnnumber{
\Dnt := \begin{pmatrix}
		\frac{1}{\alphaabcd_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaabcd_{n+1,n,1}} & \\
		\eta_{0} & 0 & \hdots & 0 & \eta_{1} & \hdots & \hdots & \eta_{n+1}
	    \end{pmatrix}, \label{eqn:ds:Dtdef}
}
 where
 \bseqn{
	\eta_{n+1} &= \frac{1}{\betaabcd_{n,n,9}}, \\ 
	\eta_{n} &= -\frac{1}{\betaabcd_{n,n-1,9}} \big( \betaabcd_{n,n,8} \: \eta_{n+1} \big), \\
	\eta_j &= - \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
	\eta_0 &= - \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
}
It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopabcd_{0}(x,y)$  through $\bighdopabcd_{n}(x,y)$ in optimal $O(n^2)$ complexity.

@@

It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in \bsrefeqn{eqn:ds:Hrecurrence}, by careful application of the Christoffel--Darboux formula \cite[(18.2.12)]{DLMF}. We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case. We thus first need to define a new set of `interim' 1D OPs.
\begin{definition}\label{def:ds:InterimOPconstruction}
Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
\bseqnnumber{
	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ds:ipgenjact}
}
where
\bseqnnumber{
	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
}
Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjact}.
\end{definition}
Note that the OPs $\{\genjac_n^{(a,b,2c)}\}$ are then equivalent to the OPs $\{\genjact_n^{(a,b,c,c)}\}$. Let the recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d)}\}$ be given by:
\bseqnnumber{
	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
}
\begin{proposition}
There exist constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ such that
\bseqnnumber{
	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansioni} \\
	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansionii}
}
\end{proposition}

\begin{proof}
Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
\bseqn{
	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
}
and define
\bseqnnumber{
	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:ds:genjactnormalisationi} \\
	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half. \label{eqn:ds:genjactnormalisationii}
}
Now, by the Christoffel--Darboux formula \cite[(18.2.12)]{DLMF}, we have that for any $x, y \in \R$,
\bseqnnumber{
	&\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) \nonumber \\
	&= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}. \label{eqn:ds:christoffeldarboux}
}
Then,
\bseqn{
	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
		&\quadfour \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
		&\quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \genjactw^{(a,b,c,d)}(x) \Big\{ \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
		&\quadeight \quadtwo \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \Big\} \: \D x \\
	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
}
using \bsrefeqn{eqn:ds:genjactnormalisationi} and \bsrefeqn{eqn:ds:christoffeldarboux}, showing that the RHS and LHS of \bsrefeqn{eqn:ds:genjacexpansioni} are equivalent. Further,
\bseqn{
	&\int_\alpha^\beta \Big\{ \big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
		&\quadfour \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big\} \: \D x \\
	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
		&\quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \genjactw^{(a,b,c,d)}(x) \Big\{ \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
		&\quadeight \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \Big\} \: \D x \\
	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
}
using \bsrefeqn{eqn:ds:genjactnormalisationii} and \bsrefeqn{eqn:ds:christoffeldarboux}, showing that the RHS and LHS of \bsrefeqn{eqn:ds:genjacexpansionii} are also equivalent. 
\end{proof}

### 2021-10-11 papers/thesis/resubmitted/ch4-spherical-caps/spherical-caps.tex
Fixed typos for Jac mats defs. Added brackets to eqn numbers in citations

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{2n+3}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bigscopta_{-1}(x,y,z) := 0 \\
		\bigscopta_{0}(x,y,z) := \scopa_0 \\
		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z)
	\end{cases}.
}

@@

It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:sc:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.

@@

\remark Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[3.5.23]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.

>>>>

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{2n+3}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p.78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bigscopta_{-1}(x,y,z) := 0 \\
		\bigscopta_{0}(x,y,z) := \scopa_0 \\
		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z)
	\end{cases}.
}

@@

It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:sc:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[(18.2.12)]{DLMF}.

@@

\remark Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[(3.5.23)]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
